\documentclass[conference]{IEEEtran}
%\documentclass[onecolumn, draft]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz,pgfplots}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Slice-and-cluster: Summarization of discrete event signals patterns with event clustering on time segments
%\thanks{Identify applicable funding agency here. If none, delete this.}
}
\author{\IEEEauthorblockN{1\textsuperscript{st} Anonymous}
\IEEEauthorblockA{\textit{Anonymous } \\
}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Anonymous}
\IEEEauthorblockA{\textit{Anonymous} \\
}
}

%\author{\IEEEauthorblockN{1\textsuperscript{st} Vanessa Cedeno-Mieles}
%\IEEEauthorblockA{\textit{Network Dynamics and Simulation Science Laboratory, } \\
%\textit{Biocomplexity Institute (of Virginia Tech)}\\
%Blacksburg, VA, USA \\
%\textit{Escuela Superior Politecnica del Litoral, ESPOL} \\
%Guayaquil, Ecuadort \\
%vcedeno@vt.edu}
%\and
%\IEEEauthorblockN{2\textsuperscript{nd} Sandeep Gupta}
%\IEEEauthorblockA{\textit{Network Dynamics and Simulation Science Laboratory, } \\
%\textit{Biocomplexity Institute (of Virginia Tech)}\\
%Blacksburg, VA, USA \\
%sandeep@vt.edu}
%}

\maketitle

\begin{abstract}
We present an approach for summarizing a given set of discrete valued time series. Our work is generalization of~\cite{Kiernan:constructing} which presents a formulation and an optimal strategy to summarize binary valued time series. 
  Summarization, concisely represents a large event sequence emphasizing the fundamental characteristics of the entire data set. Having a global overview of the dataset allows the user to explore different options of where to start a more detailed analysis. Our approach yields summaries that  1)  are concise and accurate, outputting a short and correct description of the input data. 2) provide a global data model, i.e., describe the global structure and how it evolved through the time dimension. 3) provide a local pattern model, describing local patterns so the user can identify similar behavior between events during a segment in time. and 4) is parameter-free, which means that apart from the initial event sequence, additional information need not be provided by the user.
  We use the Minimum Description Length (MDL) principle to detect regularities and compress data. We apply two dynamic programming algorithms to optimally and automatically segment the sequences in time and cluster events to discover similarities in polynomial time. Experiments on synthetic datasets effectively and efficiently summarize event sequences, finding correct cuts and clusters. Experiments on real datasets, including Ebola cases in Liberia, and the Global Landslide Catalog, demonstrate that our summarization finds useful patterns and tell stories validated with the analysis made by experts in the respective dataset domain.
\end{abstract}

\begin{IEEEkeywords}
summarization, discrete event signals, minimum description length, dynamic programming, clustering
\end{IEEEkeywords}


\input{introduction}

\input{notation}

\input{theory}

\input{model1}

\input{model2}

\input{experiments}

\input{relatedwork}

\input{conclusion}

%\section*{Acknowledgment}

%The preferred spelling of the word ``acknowledgment'' in America is without 
%an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
%G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
%acknowledgments in the unnumbered footnote on the first page.


\bibliographystyle{IEEEtran}
\bibliography{sample-bibliography}

\end{document}
